Approximately Exact Calculations for Linear Mixed Models

### 2015-06-03 ms.tex
minor

By the claim, \mrle\ must lie on or inside $B_1$, so $B_1$ could be passed to an optimizer such as \R's \textsf{optim} or \textsf{nlminb} with potentially better results than using those functions without bounds.  However, even with known bounds, general purpose optimizers may still miss \mrle\ and regions of high $\logRL$ and they emphasize single points of highest local $\logRL$ while possibly ignoring large adjacent regions where $\logRL$ is nearly as large \citep{hill:1965, hodges:2013}.

@@

\begin{algorithm}
  \caption{Learn areas in $B$ where log $f$ is high to within $\epsilon$
    \label{alg:findf}}
  \begin{algorithmic}[1]
    \Function{findf}{$B, \epsilon, M$}
      \Let{$active$}{list containing only $B$}
      \Let{$inactive$}{empty list}
      \Let{$L$}{$-\infty$}
      \While{length of $active > 0$}
        \For{each element $b$ in $active$}
          \Let{$L^b, U^b$}{get bounds on log $f$ in $b$ \Comment{see section \ref{sec:boundsinboxes}}}
        \EndFor
          \Let{tmp}{max($L^b$)}
         \Let{$L$}{max($L$, tmp)}
        \For{each element $b$ in $active$}
          \If{$U^b - L^b < \epsilon \enspace \textrm{or} \enspace L^b < L - M$}
            \State move $b$ from $active$ to $inactive$
          \Else
            \Let{$b_1, b_2, b_3, b_4$}{split $b$ into 4 smaller boxes}
            \State remove $b$ from $active$
            \State add $b_1, b_2, b_3, b_4$ to $active$
          \EndIf
        \EndFor
      \EndWhile
      \State \Return $inactive$ 
    \EndFunction
  \end{algorithmic}
\end{algorithm}

@@

Figure~\ref{fig:hmorll} depicts the same $\logRLorig$ as \cite{henn&hodges:2014}'s Figures~2a (MCMC draws) and~2b ($\logRLorig$ contours), but their Figure~2a was produced by MCMC whereas our Figure~\ref{fig:hmorll} was produced by direct calculation.  Their Figure~2a shows that the MCMC sampler did not sample any values of $\sigssq$ less than about 10, whereas our Figure~\ref{fig:hmorll} and their Figure~2b shows a region of high $\logRLorig$ extending down to $\sigssq=0$.  In fact, $\logRLorig(500,0) \approx -1241.5$, only about 6 log units below $\logRLorig(\sshat) \approx -1235.7$.  Further, about their Figure~2a, Henn and Hodges say, ``No change in contour shape indicative of a local maximum could be found in the \dots region of $(500, 600) \times (10^{-3}, 1)$, regardless of contour resolution."  I.e., they cannot be sure there are no undiscovered points with large $\logRLorig$.  In contrast, our algorithm guarantees there are no undiscovered points where $\logRLorig$ is more than $\epsilon$ above $L$.

>>>>

By the claim, \mrle\ must lie on or inside $B_1$, so $B_1$ could be passed to an optimizer such as \R's \textsf{optim} or \textsf{nlminb} with potentially better results than using those functions without bounds.  However, even with known bounds, general purpose optimizers may still miss \mrle\ and
%regions of high $\logRL$ and
they emphasize single points of highest local $\logRL$ while possibly ignoring large
%adjacent
regions where $\logRL$ is nearly as
high %large
\citep{hill:1965, hodges:2013}.

@@

\begin{algorithm}
  \caption{Learn areas in $B$ where log $f$ is high to within $\epsilon$
    \label{alg:findf}}
  \begin{algorithmic}[1]
    \Function{findf}{$B, \epsilon, M$}
      \Let{$active$}{list containing only $B$}
      \Let{$inactive$}{empty list}
      \Let{$L$}{$-\infty$}
      \While{length of $active > 0$}
        \For{each element $b$ in $active$}
          \Let{$L^b, U^b$}{get bounds on log $f$ in $b$ \Comment{see section \ref{sec:boundsinboxes}}}
        \EndFor
          \Let{tmp}{max($L^b$)}
         \Let{$L$}{max($L$, tmp)}
        \For{each element $b$ in $active$}
          \If{$U^b - L^b < \epsilon \enspace \textrm{or} \enspace U^b < L - M$}
            \State move $b$ from $active$ to $inactive$
          \Else
            \Let{$b_1, b_2, b_3, b_4$}{split $b$ into 4 smaller boxes}
            \State remove $b$ from $active$
            \State add $b_1, b_2, b_3, b_4$ to $active$
          \EndIf
        \EndFor
      \EndWhile
      \State \Return $inactive$ 
    \EndFunction
  \end{algorithmic}
\end{algorithm}

@@

Figure~\ref{fig:hmorll} depicts the same $\logRLorig$ as \cite{henn&hodges:2014}'s Figures~2a (MCMC draws) and~2b ($\logRLorig$ contours), but their Figure~2a was produced by MCMC whereas our Figure~\ref{fig:hmorll} was produced by direct calculation.  Their Figure~2a shows that the MCMC sampler did not sample any values of $\sigssq$ less than about 10, whereas our Figure~\ref{fig:hmorll} and their Figure~2b show a region of high $\logRLorig$ extending down to $\sigssq=0$.  In fact, $\logRLorig(500,0) \approx -1241.5$, only about 6 log units below $\logRLorig(\sshat) \approx -1235.7$.  Further, about their Figure~2a, Henn and Hodges say, ``No change in contour shape indicative of a local maximum could be found in the \dots region of $(500, 600) \times (10^{-3}, 1)$, regardless of contour resolution."  I.e., they cannot be sure there are no undiscovered points with large $\logRLorig$.  In contrast, our algorithm guarantees there are no undiscovered points where $\logRLorig$ is more than $\epsilon$ above $L$.
